---
Title: "Plots summarizing single-cell specific eQTL results"
Output: (1) a histogram showing how specific eQTL associations are, adn (2) scatterplots showing cis-and transregulatory regions
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Questions

These R markdown codes require the data files "c.txt" and "d.csv", provided. Codes to generate these files are provided in the two R scripts listed. Here, we show plots that summarize results for these questions:

(1) Are eQTLs (gene-expressed gene associations) specific to particular cell types or shared across cell types?

(2) Are most of these association in cis or trans-regulatory locations?

### (1) Histogram code

```{r}
dt <- read.table("c.txt", header = T)

library(ggplot2)

hist_ctype<-ggplot(dt, aes(x=Freq)) +
  geom_histogram(position="identity", alpha=0.5,bins = 10,color = 'darkgrey', fill = NA)+ #,color = 'black', fill = NA
  geom_density(alpha=0.2)+
  geom_vline(data=dt, aes(xintercept=median(Freq)),color="purple",
             linetype="dashed")+
  geom_vline(data=dt, aes(xintercept=mean(Freq),color="grey"),
             linetype="dashed", show.legend=TRUE)+
  scale_color_manual(name = "Statistics", values = c(Median = "purple", Mean = "grey"))+ #adds legend
  stat_bin(binwidth=1, geom="text", aes(label=after_stat(count)), vjust=0)+ #adds txt/freq on each bar
  scale_x_continuous(breaks = seq(1,14, by = 1)) + #adds all values of x
  labs(title="Histogram of how many cell-types a hit is found",x="Count of cell-types", y = "Frequency")+
  theme_classic() 
```

##### Plot

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(hist_ctype)
```

### (2) Scatterplot - results for each cell type on multiple pages

```{r}
d<-read.csv("d.csv", header = T)

library(ggplot2)

plotsvec<-vector("list",length=length(unique(d$cell_type)))
for (i in unique(d$cell_type)) {   #could also use levels(d$cell_type) as it is a factor
  p<-ggplot(d[d$cell_type==i,], aes(x = bp_flat, y = gbp_flat))+
    geom_pointdensity(na.rm = T) +
    scale_color_viridis(direction = 1, alpha = 0.35,option = "H")+
    geom_vline(aes(xintercept = length_flat), data = gChr_lengths,na.rm = T) +
    scale_y_continuous(breaks = axis_coordgChr,
                       labels = gChr_lengths$gChr,
                       limits = c(0, max(gChr_lengths$length_flat))) +
    scale_x_continuous(breaks = axis_coordChr,
                       labels = Chr_lengths$Chr,
                       limits = c(0, max(Chr_lengths$length_flat))) +
    ylab("Expressed Gene Chromossomic Start Location") + xlab ("eQTL Chromossomic Location") + 
    ggtitle (i)+ theme_bw()
  
  plotsvec[[i]]<-p
}
plotsvec[1:15]<-NULL #because it was adding 15 extra NULL lists int he beginning for some reason
summary(plotsvec)


```

##### Plot

```{r pressure, echo=FALSE}
plotsvec 
```

### (2) Scatterplot - results for each cell type on a single panel

```{r}

# Facet wrap####
library(ggplot2)
p<-ggplot(d, aes(x = bp_flat, y = gbp_flat))+
    facet_wrap(cell_type ~ .,scales = "free", ncol = 3)+ #this separate plots by cell types, facet_wrap is used for a variable with many levels
    geom_pointdensity() +
    scale_color_viridis(direction = 1, alpha = 0.35,option = "H")+
    geom_vline(aes(xintercept = length_flat), data = gChr_lengths) +
    scale_y_continuous(breaks = axis_coordgChr,
                       labels = gChr_lengths$gChr,
                       limits = c(0, max(gChr_lengths$length_flat))) +
    scale_x_continuous(breaks = axis_coordChr,
                       labels = Chr_lengths$Chr,
                       limits = c(0, max(Chr_lengths$length_flat))) +
    ylab("Expressed Gene Chromossomic Start Location") + xlab ("eQTL Chromossomic Location") + 
    ggtitle ("Cis-trans regulatory locations by cell-type")+ theme_bw()
  
 # Density of associations across cell types (visualysing the count (density) of dots inside hexagonal areas)
p2<-ggplot(d, aes(x = bp_flat, y = gbp_flat))+
    facet_wrap(cell_type ~ .,scales = "free", ncol = 3)+ #this separate plots by cell types, facet_wrap is used for a variable with many levels
  geom_hex (bins=3, aes(alpha= ..count..))+
  # scale_color_viridis(direction = 1, alpha = 0.35,option = "H")+
    geom_vline(aes(xintercept = length_flat), data = gChr_lengths) +
    scale_y_continuous(breaks = axis_coordgChr,
                       labels = gChr_lengths$gChr,
                       limits = c(0, max(gChr_lengths$length_flat))) +
    scale_x_continuous(breaks = axis_coordChr,
                       labels = Chr_lengths$Chr,
                       limits = c(0, max(Chr_lengths$length_flat))) +
    ylab("Expressed Gene Chromossomic Start Location") + xlab ("eQTL Chromossomic Location") + 
    ggtitle ("Cis-trans regulatory locations by cell-type")+ theme_bw()

```

```{r pressure, echo=FALSE}
#Facet wrap - within cell-type patters
plot(p) 

```

```{r pressure, echo=FALSE}
#Density of dots across all cell types
plot(p2)
```
